---
title: "HMK 8: summarising, reshaping, merging"
format: gfm
editor: visual
---

```{r message = FALSE}
library(tidyverse)
library(nycflights13)
```

# Q1: Summarizing operations and exploratory data analysis

## Q1a

Download `experiment1.csv` from Canvas (under files). This reports two variables from an experiment that has four treatments. Creatively, the treatments are `1`, `2`, `3`, and `4`, and the two variables are `x` and `y`.

Create a data frame that contains the mean, standard deviation, and number of points for each treatment. You will want to use functions like `summarise()`, `group_by()`, `mean()`, and `sd()`.

Importing data

```{r}
experiment1 <- read.csv("experiment1.csv")
```

Below I use the `group_by` function so I can analyze specific groups rather than the entire data set when I use the `summarise` function.

I create new columns with informative names indicating the performed functions `mean` and `sd` for variables `x` and `y`.

```{r}
experiment1_stats <- group_by(experiment1, balls)
summarise(experiment1_stats, mean_x = mean(x), mean_y = mean(y),
          sd_x = sd(x), sd_y = sd(y))
```

Are the data sets different in any important way?

The treatments are not necessarily different from each other. The only discernible difference is that the standard deviation of the `x` variable is slightly larger than `y`'s.

## Q1b

Now load the file `experiment2.csv`. Again, this describes two variables for multiple treatments (here called `dataset`). Answer the same questions as above.

Importing data:

```{r}
experiment2 <- read.csv("experiment2.csv")
```

Carrying out same steps as Q1a:

```{r}
experiment2_stats <- group_by(experiment2, dataset)
summarise(experiment2_stats, mean_x = mean(x), mean_y = mean(y),
          sd_x = sd(x), sd_y = sd(y))
```

Are the datasets different in any important way?

It would not appear that there is any major difference between these data sets. Variable `x` has a smaller standard deviation than variable `y`.

# Q2: pivoting

Create a plot that illustrates the differences in income among religions, using the `relig_income` data set that is built into tidyverse.

You will need to create a tidy data frame to do this.

```{r}
# adding relig_income to my environment so I can look at it easier
relig_income <- relig_income
```

```{r}
# using pivot longer and the dplyer::select notation to pivot multiple columns
relig_income_tidy <- relig_income %>% 
  pivot_longer(!religion, names_to = "income", values_to = "count")  %>%
  group_by(religion)

relig_income_tidy$income <- factor(relig_income_tidy$income,
 levels = c("<$10k", "$10-20k", "$20-30k", "$30-40k", "$40-50k", "$50-75k", "$75-100k", "$100-150k", ">150k", "Don't know/refused"))
```

```{r}
ggplot(relig_income_tidy, aes(x = income, y = count, fill = religion)) +
  geom_col() +
  scale_x_discrete(guide = guide_axis(n.dodge=2)) +
  labs(x = "income range", y = "count", fill = "religion") +
  guides(fill = guide_legend(ncol=4)) +
  ggtitle("Religions vs Income Range") +
  theme(legend.position = "bottom", legend.key.size = unit(0.8, "lines"))
# I used `theme_update(plot.title = element_text(hjust = 0.5))` in the 
# command line so future plots will have a centered title.
```

# Q3: merging

## Q3a: meaning of joins

Explain the difference between a left join, a right join, an inner join, and an outer join.

Outer join: `left`, `right` and `full` joins are all types of `outer joins`

Left join: Preserves observations from the `x` table, should be preferred join as it keeps original observations if there is not a match in the `y` table.

Right join: Similar to `left join` but preserves observations from the second or `y` table.

Full join: Keeps both `x` and `y` observations

Inner join: A join that matches pairs of equal observations

## Q3b: using joins

Using the `flights` and `weather` data sets from `nycflights13`, determine whether there is a correlation between average hourly wind speed and departure delays at NY airports.

This is a question about joins: you will need to join the `flights` and `weather` by year, month, day, and hour. However, note that `flights` has encoded departure time in a particularly annoying way: as an integers. For instance, the integer 517 indicates 5:17 am.

`weather` gives average weather conditions each hour, with the hour given as an integer (e.g., 5 indicating 5-6 am). You're going to have to figure out how to convert the time in `flights` into a form that matches the form in `weather`.

If you want, you can use the `lm()` function to make a linear model of departure delay as a function of wind speed. But it is also fine to just make a plot of the two variables with `geom_smooth()`.

```{r}

# assinging datasets to objects for ease of access
f <- flights
w <- weather

# dividing `dep_time` by 100 using integer division
f <- f %>%
  mutate(f, hour = dep_time %/% 100) %>%
  select(origin, year, month, day, hour, dep_delay)
```

```{r}
weather_grouped <- select(weather, year, month, day, hour, wind_speed)
```

```{r}
wf <- weather_grouped %>%
left_join(f, weather_grouped, by = c("year", "month", "day", "hour"))
```

```{r}
ggplot(wf, aes(x = wind_speed, y = dep_delay)) +
  geom_smooth(method = "lm")
```
