---
title: "hmk_02"
author: "murray stokes"
format: gfm
editor: visual
---

### Load tidyverse and suppress its message

```{r message = FALSE}
library(tidyverse)
```

### Assign values to objects "a" and "b"

```{r}
a <- 4/9
b <- 4/11
```

### Determine if a is less than b

```{r}
a > b
```

### List objects in environment

```{r}
ls()
```

### Remove objects from environment

```{r}
rm(list = ls())
```

The syntax for clearing the environment is structured in this "weird" way because the first argument that is "seen" by rm(), is "..." rather than "list". If "list" was the first (default) argument, then simply typing rm(ls()) would be sufficient in clearing the environment. Because of this, we must indicate to rm() that we would like it to use the list argument to name the objects to remove ( the objects in our environment).

### List objects in environment

```{r}
ls()
```

### Display loaded packages

```{r}
search()
```

This search function shows the order that R finds objects.

For example, R will look in .GlobalEnv, before tools:rstudio
